---
title: "AS! Gaitspeed Analyses"
author: "Ryan Stanley Falck"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Set-up

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(boot,MASS,readxl,mediation,psych,plyr,sandwich,ggplot2,lme4,lmerTest,rpart,survival,survminer,lsmeans,fmsb, mice, mitools, tableone)

setwd("~/Desktop/Manuscripts/Ongoing Projects/ActionSeniors! Cognitive Frailty/")
data<-read_excel("ActionSeniors_CogsSPPB August 24, 2023.xlsx")#Data#

data$exclude<-0
data$exclude[data$ID==330]<-1

data$Group.1<-NA
data$Group.1[data$Group==0]<-"Interview"
data$Group.1[data$Group==1]<-"Otago"

data$Group_r[data$Group==0]<-1
data$Group_r[data$Group==1]<-0

data$Gender<-NA
data$Gender[data$Sex==1]<-0
data$Gender[data$Sex==2]<-1

data$Sex[data$Sex==1]<-"Female"
data$Sex[data$Sex==2]<-"Male"

#Create long data set
library(plyr)
wide<-rename(data,c("Weight_1"="Weightbaseline", "Height_1"="Heightbaseline","BMI_1"="BMIbaseline",
                    "Falls 12M prior to baseline"="Fallsbeforebaseline", "Total_Exposure"="TotalExposure",
                    "Total_Falls"="TotalFalls","PPA_1"="PPAbaseline","GDS_1"="GDSbaseline","MoCA_1"="MoCAbaseline",
                    "Trails_BA_1"="TrailsBAbaseline","DSST_1"="DSSTbaseline","TUG_1"="TUGbaseline",
                    "SPPB_1"="SPPBbaseline","Gait_Speed_1"="GaitSpeedbaseline", "Digits_F_1"="DigitsFbaseline",
                    "Digits_B_1"="DigitsBbaseline", "Stroop_3_1"="Stroopbaseline", "Trails_BA_2"="TrailsBA_2",
                    "Trails_BA_3"="TrailsBA_3","Group_r"="Group.r","Gait_Speed_2"="GaitSpeed_2","Gait_Speed_3"="GaitSpeed_3",
                    "Total Falls 6 Months"="TotalFalls6Months", "Stroop_3_2"="Stroop_2", "Stroop_3_3"="Stroop_3",
                    "Digits_F_2"="DigitsF_2","Digits_B_2"="DigitsB_2","Digits_F_3"="DigitsF_3","Digits_B_3"="DigitsB_3"))
detach("package:plyr", unload = TRUE)

wide$time<-1

wide2<-wide[c(1,64:70,2:18,21,24:29,31:34,37,40:45,47:50,53,56:61,63)]
data2 <- reshape(as.data.frame(wide2),idvar="ID",varying=c(34:55),direction="long",sep="_")

data2$Stroopbaseline<-as.numeric(data2$Stroopbaseline)
data2$Stroop<-as.numeric(data2$Stroop)
data2$TUGbaseline<-as.numeric(data2$TUGbaseline)
data2$TUG<-as.numeric(data2$TUG)
data2$timefactor<-as.factor(data2$time)

wide2$Stroopbaseline<-as.numeric(wide2$Stroopbaseline)
wide2$TUGbaseline<-as.numeric(wide2$TUGbaseline)
wide2$Stroop_2<-as.numeric(wide2$Stroop_2)
wide2$Stroop_3<-as.numeric(wide2$Stroop_3)
```

# 2) Binary variable of slow gait speed

Participants categorized slow (i.e., below mean gait speed of 0.87 m/s) vs. fast gait speed at baseline
```{r}
data2$slowgait<-NA
data2$slowgait[data2$GaitSpeedbaseline<=0.8]<-0
data2$slowgait[data2$GaitSpeedbaseline>0.8]<-1

wide2$slowgait<-NA
wide2$slowgait[wide2$GaitSpeedbaseline<=0.8]<-0
wide2$slowgait[wide2$GaitSpeedbaseline>0.8]<-1
```

# 3) Figure 1 - CONSORT Flow Chart Stratified by Baseline Gait Speed Status

```{r}
#Determine CogFrail Status of Dropouts
CONSORT<- subset(wide2, ID==128 | ID== 131 | ID==132 | ID==134 | ID==158 | ID==163 | ID==176 | ID==185 | ID==192 | ID==195 | ID==213 | ID==214 | ID==216 | ID==220 | ID==222 | ID==223 |
                   ID==270 | ID==278 | ID==281 | ID==284 | ID==306 | ID==310 | ID==318 | ID==321 | ID==330 | ID==334 | ID==339 | ID==357 | ID==362 | ID==369 | ID==373 | ID==375 | ID==378 |
                   ID==385 | ID==395 | ID==398 | ID==407 | ID==408 | ID==412 | ID==413 | ID==424 | ID==426 | ID==427 | ID==428 | ID==432 | ID==438 | ID==455 | ID==461) #All dropouts

CONSORT<-CONSORT[c(1,9,56)]
MONTHS<- c(13,2,1,0,0,0,13,0,0,13,2,4,0,11,3,5,1,5,2,5,0,0,3,0,0,5,0,4,5,5,5,7,6,1,1,9,1,7,9,2,4,1,7,7,5,11,7,10) #Months in study before drop-out
CONSORT<-cbind(CONSORT,MONTHS) 

CONSORT$Group[CONSORT$Group==1]<-"INT"
CONSORT$Group[CONSORT$Group==0]<-"CON"

#Dropouts before 6 months (i.e., Month 7) or after 6 months in study
CONSORT$Timepoint<-NA
CONSORT$Timepoint[CONSORT$MONTHS<=7]<-"6 Months or Less"
CONSORT$Timepoint[CONSORT$MONTHS>7]<-"More than 6 Months"


#Sort by Group, Gait Speed Status, and Time of Dropout
table(CONSORT$Timepoint,CONSORT$Group, CONSORT$slowgait)
```

# 4) Baseline characteristics

Stratified datasets
```{r}
fastgait.wide<-subset(wide2, slowgait==1)
slowgait.wide<-subset(wide2, slowgait==0)

fastgait.long<-subset(data2, slowgait==1)
slowgait.long<-subset(data2, slowgait==0)
```

Baseline variables
```{r}
baseline<-dput(names(wide2[c(11,10,14,15,16,21:23,20,31,32,33,24,25,30,26,27)]))
```

## 4.1) All participants

```{r}
Table1_full<-CreateTableOne(vars=baseline, strata="Group", data=wide2)
print(Table1_full,contDigits=2,missing=TRUE,quote=TRUE)
```

## 4.2) Participants stratified by gait speed classification
```{r}
Table1_full<-CreateTableOne(vars=baseline, strata="slowgait", data=wide2)
print(Table1_full,contDigits=2,missing=TRUE,quote=TRUE)
```

# 4.3) Slow gait only

```{r}
Table1_slow<-CreateTableOne(vars=baseline, strata="Group", data=slowgait.wide)
print(Table1_slow,contDigits=2,missing=TRUE,quote=TRUE)
```

# 4.4) Fast gait only

```{r}
Table1_fast<-CreateTableOne(vars=baseline, strata="Group", data=fastgait.wide)
print(Table1_fast,contDigits=2,missing=TRUE,quote=TRUE)
```


# 5) Primary outcome models from JAMA paper (Liu-Ambrose et al., 2019)

Falls risk over 12 months
```{r}
mdl.nb<-glm.nb(TotalFalls~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(data2,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
print("Significant effect of intervention on falls risk at 12 months for participants with CogFrailty")  
```

Falls risk over 6 months
```{r}
mdl2.nb<-glm.nb(TotalFalls6Months~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/182.5)),subset(data2,exclude==0)) 
print(summary(mdl2.nb))
SE<-sqrt(diag(vcovHC(mdl2.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl2.nb)/SE))
print(cbind(coef(mdl2.nb),SE,p))
est<-cbind(Estimate=coef(mdl2.nb),LL=coef(mdl2.nb)-1.96*SE,UL=coef(mdl2.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
print("Significant effect of intervention on falls risk at 6 months for participants with CogFrailty")    
```


# 6) Effects of the intervention for slow vs. fast gait speed

## 6.1) Interaction effects

Falls over 12 months
```{r}
mdl.nb<-glm.nb(TotalFalls~Group.1*slowgait+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(wide2,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```

Falls over 6 months
```{r}
mdl.nb<-glm.nb(TotalFalls6Months~Group.1*slowgait+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(wide2,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```

## 6.2) Fast gait only

Falls over 12 months
```{r}
mdl.nb<-glm.nb(TotalFalls~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(fastgait.wide,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```

Falls over 6 months
```{r}
mdl.nb<-glm.nb(TotalFalls6Months~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(fastgait.wide,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```


## 6.3) Slow gait only

Falls over 12 months
```{r}
mdl.nb<-glm.nb(TotalFalls~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(slowgait.wide,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```

Falls over 6 months
```{r}
mdl.nb<-glm.nb(TotalFalls6Months~Group.1+scale(Gender,scale=FALSE)+offset(log(TotalExposure/365)),subset(slowgait.wide,exclude==0)) 
print(summary(mdl.nb))
SE<-sqrt(diag(vcovHC(mdl.nb,"HC1")))
p=2*pnorm(-abs(coef(mdl.nb)/SE))
print(cbind(coef(mdl.nb),SE,p))
est<-cbind(Estimate=coef(mdl.nb),LL=coef(mdl.nb)-1.96*SE,UL=coef(mdl.nb)+1.96*SE)
exp_group<-exp(est)
print("Use these estimates and LL and UL for the 95% CI");exp_group
```


## 6.4) Post-hoc contrasts

Fast vs. Slow at 12 months
```{r}
z=(-0.40 - (-0.46))/(sqrt((0.21)^2 + (0.25)^2))
p=2*pnorm(-abs(z))

z
p
```

Fast vs. Slow at 6 months
```{r}
z=(-0.13 - (-0.57))/(sqrt((0.23)^2 + (0.27)^2))
p=2*pnorm(-abs(z))

z
p
```

## 6.5) Cumulative Falls by Intervention Group and Gait Speed

Set-up for cumulative falls
```{r}
setwd("~/Desktop/Manuscripts/Ongoing Projects/ActionSeniors! Cognitive Frailty/")
Monthly_falls <- read_excel("AS_Monthly falls_Mar 20 2019.xlsx")
Monthly_falls<-subset(Monthly_falls, ID %in% wide2$ID)

data1c<-cbind(Monthly_falls,wide2$slowgait,wide2$Group)
data1c<-subset(data1c,ID!=330)

data1c<-plyr::rename(data1c,c("Month_1_Cumulative"="MonthCumulative_1","Month_2_Cumulative"="MonthCumulative_2","Month_3_Cumulative"="MonthCumulative_3",
                              "Month_4_Cumulative"="MonthCumulative_4","Month_5_Cumulative"="MonthCumulative_5","Month_6_Cumulative"="MonthCumulative_6",
                              "Month_7_Cumulative"="MonthCumulative_7","Month_8_Cumulative"="MonthCumulative_8","Month_9_Cumulative"="MonthCumulative_9",
                              "Month_10_Cumulative"="MonthCumulative_10","Month_11_Cumulative"="MonthCumulative_11","Month_12_Cumulative"="MonthCumulative_12",
                              "Month_13_Cumulative"="MonthCumulative_13","wide2$slowgait"="slowgait","wide2$Group"="Group"))
data1c_normalgait_EX<-subset(data1c,Group==1 & slowgait==1)
data1c_normalgait_UC<-subset(data1c,Group==0 & slowgait==1)
data1c_slowgait_EX<-subset(data1c,Group==1 & slowgait==0)
data1c_slowgait_UC<-subset(data1c,Group==0 & slowgait==0)

print("Normal Gait, Exercise");apply(data1c_normalgait_EX[c(2:14)], 2, function(x) length(which(!is.na(x)))) #count people with falls data
print("Normal Gait, Usual Care");apply(data1c_normalgait_UC[c(2:14)], 2, function(x) length(which(!is.na(x)))) #count people with falls data
print("Slow Gait, Exercise");apply(data1c_slowgait_EX[c(2:14)], 2, function(x) length(which(!is.na(x)))) #count people with falls data
print("Slow Gait, Usual Care");apply(data1c_slowgait_UC[c(2:14)], 2, function(x) length(which(!is.na(x)))) #count people with falls data

data1c.stacked<-reshape(data1c,idvar="ID",varying=c(2:27),direction="long",sep="_")
data1c.stacked$time<-data1c.stacked$time-1

Fallsmeans<-data.frame(aggregate(MonthCumulative~Group*slowgait+time,data1c.stacked,sum))
subset(Fallsmeans, slowgait==0)
```

Figure for accumulation of falls by treatment group for participants with slow gait speed

```{r}
Fallsmeans$SlowgaitGroup<-NA
Fallsmeans$SlowgaitGroup[Fallsmeans$Group==0 & Fallsmeans$slowgait==0]<-"Usual Care with Slow Gait Speed"
Fallsmeans$SlowgaitGroup[Fallsmeans$Group==1 & Fallsmeans$slowgait==0]<-"Exercise with Slow Gait Speed"
Fallsmeans$SlowgaitGroup[Fallsmeans$Group==0 & Fallsmeans$slowgait==1]<-"Usual Care with Normal Gait Speed"
Fallsmeans$SlowgaitGroup[Fallsmeans$Group==1 & Fallsmeans$slowgait==1]<-"Exercise with Normal Gait Speed"

Fallsmeans.cogfrail<-subset(Fallsmeans)

Cumulativefalls.Graph<- Fallsmeans.cogfrail %>%
  ggplot() + geom_line(aes(x=time, y=MonthCumulative, color=SlowgaitGroup)) +
  labs(x="Months", y="Cumulative number of falls") + labs(color="Intervention Group") +
  scale_x_continuous(breaks=seq(0,13,by=1)) + scale_y_continuous(breaks=seq(0,300,by=50))+
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
Cumulativefalls.Graph
```


## 6.6) Estimated differences in falls per person-year at 12 months

```{r}
wide2$falls_ppy<-wide2$TotalFalls/(wide2$TotalExposure/365)

fallsppy_lm<-lm(falls_ppy~Group*slowgait + scale(Gender,scale=FALSE),  subset(wide2, exclude==0))
anova(fallsppy_lm)
lsmeans(fallsppy_lm, ~Group|slowgait)
contrast(lsmeans(fallsppy_lm, ~~Group|slowgait), "trt.vs.ctrl", adj="none")
```

## 6.6.1) Post-hoc

Matrix
```{r}
#EX vs. usual care differences at 6 months between slow gait and normal gait
contrast1 = matrix(c(-1,1,0,0))
contrast2 = matrix(c(0,0,-1,1))
contrast_result = contrast1 - contrast2
```

Post-hoc contrasts
```{r}
contrast(lsmeans(fallsppy_lm, ~~Group|slowgait),list(contrast_result),by=NULL)
```

## 6.7) Estimated differences in falls per person-year at 6 months

```{r}
wide2$falls_ppy_6mnths<-wide2$TotalFalls6Months/(wide2$TotalExposure/182.5)

fallsppy_6mnths_lm<-lm(falls_ppy_6mnths~Group*slowgait + scale(Gender,scale=FALSE),  subset(wide2, exclude==0))
anova(fallsppy_6mnths_lm)
lsmeans(fallsppy_6mnths_lm, ~Group|slowgait)
contrast(lsmeans(fallsppy_6mnths_lm, ~~Group|slowgait), "trt.vs.ctrl", adj="none")
```

### 6.7.1) Post-hoc

Post-hoc contrasts
```{r}
contrast(lsmeans(fallsppy_6mnths_lm, ~~Group|slowgait),list(contrast_result),by=NULL)
```

## 6.8) Differences in exposure time

```{r}
expopsure_lm<-lm(TotalExposure~Group*slowgait,  subset(wide2, exclude==0))
anova(expopsure_lm)
lsmeans(expopsure_lm, ~Group|slowgait)
contrast(lsmeans(expopsure_lm, ~~Group|slowgait), "trt.vs.ctrl", adj="none")
```



# 7) Effects of the intervention on gait speed

Main effects
```{r}
Gaitspeed_lm<-lmer(GaitSpeed~Group.1*timefactor + Gender + GaitSpeedbaseline + (1|ID),  subset(data2, exclude==0))
anova(Gaitspeed_lm)
lsmeans(Gaitspeed_lm, ~Group.1|timefactor)
contrast(lsmeans(Gaitspeed_lm, ~Group.1|timefactor), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(Gaitspeed_lm, ~Group.1|timefactor), "trt.vs.ctrl", adj="none"))
```


Differences in effects for fast vs.slow gait speed
```{r}
Gaitspeed_lm_mod<-lmer(GaitSpeed~Group.1*timefactor*slowgait + Gender + GaitSpeedbaseline + (1|ID),  subset(data2, exclude==0))
anova(Gaitspeed_lm_mod)
lsmeans(Gaitspeed_lm_mod, ~Group.1|timefactor|slowgait)
contrast(lsmeans(Gaitspeed_lm_mod, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(Gaitspeed_lm_mod, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```


# 8) Does baseline gait speed modify intervention effects on secondary outcomes?

Posthoc Contrast Statements
```{r}
#EX vs. usual care differences at 6 months between slow gait and normal gait
c1 = matrix(c(0,1,0,0,0,-1,0,0))
c2 = matrix(c(1,0,0,0,-1,0,0,0))
c1st = c1 - c2

#EX vs. usual care differences at 12 months between slow gait and normal gait
c3 = matrix(c(0,0,0,1,0,0,0,-1))
c4 = matrix(c(0,0,1,0,0,0,-1,0))
c2nd = c3 - c4
```

## 8.1) SPPB
```{r}
SPPB_lm<-lmer(SPPB~Group.1*timefactor*slowgait + Gender + SPPBbaseline + (1|ID),  subset(data2, exclude==0))
anova(SPPB_lm)
lsmeans(SPPB_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(SPPB_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(SPPB_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

## 8.2) PPA
```{r}
PPA_lm<-lmer(PPA~Group.1*timefactor*slowgait + Gender + PPAbaseline + (1|ID),  subset(data2, exclude==0))
anova(PPA_lm)
lsmeans(PPA_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(PPA_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(PPA_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

## 8.3) TUG
```{r}
TUG_lm<-lmer(TUG~Group.1*timefactor*slowgait + Gender + TUGbaseline + (1|ID),  subset(data2, exclude==0))
anova(TUG_lm)
lsmeans(TUG_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(TUG_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(TUG_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

Post-hoc contrasts
```{r}
contrast(lsmeans(TUG_lm, ~Group.1|timefactor|slowgait),list(c1st,c2nd),by=NULL)
```

## 8.4) MoCA
```{r}
MoCA_lm<-lmer(MoCA~Group.1*timefactor*slowgait + Gender + MoCAbaseline + (1|ID),  subset(data2, exclude==0))
anova(MoCA_lm)
lsmeans(MoCA_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(MoCA_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(MoCA_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

## 8.5) DSST
```{r}
DSST_lm<-lmer(DSST~Group.1*timefactor*slowgait + Gender + DSSTbaseline + (1|ID),  subset(data2, exclude==0))
anova(DSST_lm)
lsmeans(DSST_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(DSST_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(DSST_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

```{r}
contrast(lsmeans(DSST_lm, ~Group.1|timefactor|slowgait),list(c1st,c2nd),by=NULL)
```

## 8.6) Trails
```{r}
Trails_lm<-lmer(TrailsBA~Group.1*timefactor*slowgait + Gender + TrailsBAbaseline + (1|ID),  subset(data2, exclude==0))
anova(Trails_lm)
lsmeans(Trails_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(Trails_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(Trails_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```

## 8.7) Stroop
```{r}
Stroop_lm<-lmer(Stroop~Group.1*timefactor*slowgait + Gender + Stroopbaseline + (1|ID),  subset(data2, exclude==0))
anova(Stroop_lm)
lsmeans(Stroop_lm, ~Group.1|timefactor|slowgait)
contrast(lsmeans(Stroop_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none")
confint(contrast(lsmeans(Stroop_lm, ~Group.1|timefactor|slowgait), "trt.vs.ctrl", adj="none"))
```


# 9) Compliance data

## 9.1) Compliance data set-up
```{r}
setwd("~/Desktop/Manuscripts/Ongoing Projects/ActionSeniors! Cognitive Frailty/")
compliancedata<-read_excel("AS_compliance_August 24, 2023.xlsx") #Data#
library(plyr)
compliancedata2<-rename(compliancedata, c("Month 1"="OEPCOMP_1","Month 2"="OEPCOMP_2","Month 3"="OEPCOMP_3","Month 4"="OEPCOMP_4","Month 5"="OEPCOMP_5","Month 6"="OEPCOMP_6",
                                          "Month 7"="OEPCOMP_7","Month 8"="OEPCOMP_8","Month 9"="OEPCOMP_9","Month 10"="OEPCOMP_10","Month 11"="OEPCOMP_11","Month 12"="OEPCOMP_12"))
detach("package:plyr", unload = TRUE)

#Maximize compliance at 100% for each individual month
compliancedata2$OEPCOMP_1[compliancedata2$OEPCOMP_1>100]<-100
compliancedata2$OEPCOMP_2[compliancedata2$OEPCOMP_2>100]<-100
compliancedata2$OEPCOMP_3[compliancedata2$OEPCOMP_3>100]<-100
compliancedata2$OEPCOMP_4[compliancedata2$OEPCOMP_4>100]<-100
compliancedata2$OEPCOMP_5[compliancedata2$OEPCOMP_5>100]<-100
compliancedata2$OEPCOMP_6[compliancedata2$OEPCOMP_6>100]<-100
compliancedata2$OEPCOMP_7[compliancedata2$OEPCOMP_7>100]<-100
compliancedata2$OEPCOMP_8[compliancedata2$OEPCOMP_8>100]<-100
compliancedata2$OEPCOMP_9[compliancedata2$OEPCOMP_9>100]<-100
compliancedata2$OEPCOMP_10[compliancedata2$OEPCOMP_10>100]<-100
compliancedata2$OEPCOMP_11[compliancedata2$OEPCOMP_11>100]<-100
compliancedata2$OEPCOMP_12[compliancedata2$OEPCOMP_12>100]<-100


compliancedata2$AverageComply<-(compliancedata2[c(2)] + compliancedata2[c(3)] + compliancedata2[c(4)] + compliancedata2[c(5)] + compliancedata2[c(6)] + compliancedata2[c(7)] + 
                                  compliancedata2[c(8)] + compliancedata2[c(9)] + compliancedata2[c(10)] + compliancedata2[c(11)] + compliancedata2[c(12)] + compliancedata2[c(13)])/12

wide3<- merge(compliancedata2,wide2, by = "ID")
wide3$slowgait[wide3$slowgait==1]<-"Normal Gait Speed"
wide3$slowgait[wide3$slowgait==0]<-"Slow Gait Speed"

complygraph<-wide3[c(1:13,69)]

#long form
compliancelong <- reshape(as.data.frame(complygraph),idvar="ID",varying=c(2:13),direction="long",sep="_")

#Maximize compliance at 100% (i.e., compliance cannot be greater than 100%, regardless of how many times/month OEP was completed)
compliancelong$OEPCOMP[compliancelong$OEPCOMP>100]<-100


compliancelong$Timepoint<-as.factor(compliancelong$time)
```

## 9.2) Average compliance by gait speed
```{r}

describeBy(wide3$AverageComply, wide3$slowgait)
```

## 9.3) Compliance data by month
```{r}
comply.2<-lm(OEPCOMP~factor(time)*slowgait, compliancelong)
emms.2<-as.data.frame(lsmeans(comply.2, ~factor(time)|slowgait))
print(emms.2)
```

## 9.4) Compliance data graph
```{r}
Compliance.Graph<- ggplot(emms.2, aes(x=time, y=lsmean, group = slowgait)) + geom_line(aes(linetype= slowgait)) +
  labs(x="Months", y="Monthy Adherence %") + ylim(0,100) + scale_x_continuous(breaks=seq(0,13,by=2)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
Compliance.Graph
```


# 10) Additional Analyses for Response to Reviewers


## 10.1) Comparing attrition rates between groups

Set-up data frame comparing attrition rates between groups
```{r}
Attrition<-wide2[c(1,5,56)]

library(dplyr)
Attrition2 <- Attrition %>%
  mutate(Dropouts = ifelse(ID==128 | ID== 131 | ID==132 | ID==134 | ID==158 | ID==163 | ID==176 | ID==185 | ID==192 | ID==195 |
                                ID==213 | ID==214 | ID==216 | ID==220 | ID==222 | ID==223 | ID==270 | ID==278 | ID==281 | ID==284 |
                                ID==306 | ID==310 | ID==318 | ID==321 | ID==330 | ID==334 | ID==339 | ID==357 | ID==362 | ID==369 |
                                ID==373 | ID==375 | ID==378 | ID==385 | ID==395 | ID==398 | ID==407 | ID==408 | ID==412 | ID==413 |
                                ID==424 | ID==426 | ID==427 | ID==428 | ID==432 | ID==438 | ID==455 | ID==461, "Dropout","Completed"),
         Dropout_6 = ifelse(Dropouts=="Dropout" & (ID== 131 | ID==132 | ID==134 | ID==158 | ID==163 | ID==185 | ID==192 | ID==213 |
                                                       ID==214 | ID==216 | ID==222 | ID==223 | ID==270 | ID==278 | ID==281 | ID==284 |
                                ID==306 | ID==310 | ID==318 | ID==321 | ID==330 | ID==334 | ID==339 | ID==357 | ID==362 | ID==369 |
                                ID==373 | ID==375 | ID==378 | ID==385 | ID==395 | ID==407 | ID==408 | ID==413 | ID==424 | ID==426 | 
                                  ID==427 | ID==428 | ID==432 | ID==455), 1, NA),
         Dropout_12 = ifelse(Dropouts=="Dropout" & (ID==128 | ID==176 | ID==195 | ID==220 | ID==398 | ID==412 | ID==438 | ID==461), 1, NA)
           )
```


```{r}
table(Attrition2$Group.1, Attrition2$Dropouts, Attrition2$slowgait)

table_data <- xtabs(~ Group.1 + Dropouts + slowgait, data = Attrition2)

mantelhaen.test(table_data)
```


## 10.2) Simplified chi-square for fall rates based on gait speed categorization

```{r}
EX_slow<- subset(data1c, Group == 1 & slowgait == 1)
UC_slow<- subset(data1c, Group == 0 & slowgait == 1)
EX_fast<- subset(data1c, Group == 1 & slowgait == 0)
UC_fast<- subset(data1c, Group == 0 & slowgait == 0)

Total_falls_EX_slow <- sum(EX_slow$MonthCumulative_13, na.rm = TRUE)
Total_falls_UC_slow <- sum(UC_slow$MonthCumulative_13, na.rm = TRUE)
Total_falls_EX_fast <- sum(EX_fast$MonthCumulative_13, na.rm = TRUE)
Total_falls_UC_fast <- sum(UC_fast$MonthCumulative_13, na.rm = TRUE)

Total_falls_EX_slow
Total_falls_UC_slow
Total_falls_EX_fast
Total_falls_UC_fast

group <- c("Interview", "Otago")
gaitspeed <- c("Slow", "Normal")

table_data2 <-c(196,132,99,168)

dim_data <- c(2,2)

matrix_array <-array(table_data2, dim = dim_data,
                     dimnames = list(Group = group, Gaitspeed = gaitspeed))
#Total Falls
print(matrix_array)

if (!require("DescTools", character.only = TRUE)) {
  install.packages("DescTools")
  library("DescTools", character.only = TRUE)
}

result <- GTest(matrix_array)

print(result)


colnames(wide2)

# Fit the model without interaction (null model)
null_model <- glm(TotalFalls ~ Group + slowgait + Gender, data = wide2, family = poisson)


# Fit the model with interaction (full model)
full_model <- glm(TotalFalls ~ Group * slowgait + Gender, data = wide2, family = poisson)

# Perform the likelihood ratio test
lrt_result <- anova(null_model, full_model, test = "LRT")
print(lrt_result)
```

## 10.3) Estimating group differences in TUG at 6 months or 12 months based on the criterion of 13.5 sec
```{r}
wide4<-wide2 %>%
  mutate(
    TUGbaseline_cri=case_when(TUGbaseline>=13.5~1,
                        TUGbaseline<13.5~0),
    TUG_2_cri=case_when(TUG_2>=13.5~1,
                        TUG_2<13.5~0),
    TUG_3_cri=case_when(TUG_3>=13.5~1,
                        TUG_3<13.5~0)
  )

log.TUG2<-glm(TUG_2_cri~Group*slowgait + Gender + TUGbaseline_cri, data=wide4, family = "binomial")
summary(log.TUG2)
exp(coef(log.TUG2))
exp(cbind(OR = coef(log.TUG2), confint(log.TUG2)))

log.TUG3<-lm(TUG_3_cri~Group*slowgait + Gender + TUGbaseline_cri, data=wide4, family = "binomial")
summary(log.TUG3)
exp(coef(log.TUG3))
exp(cbind(OR = coef(log.TUG3), confint(log.TUG3)))
```